---
title: "Base plotting system"
author: "Yuling"
date: "4/13/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Core plotting and graphics engine packages: 
* graphisc
* grDevices

# The lattice plotting system packages:
* lattice
* grid

# The process of making a plot
* Where the plot be made decides the quality of the plot;
* What type of plot
* Do you need to dynamically resize the graphic;

# Base Graphics
* Mostly useful for 2-D graphics
* There are two phases to creating a base plot  
  -- Initializing a new plot  
  -- Annotating (adding to) an existing plot  

* Calling plot(x, y) or hist(x,y) will launch a graphics device
  + ?plot to see all the arguments for setting the title, x axis label, y axis label, etc.
  + ?par, parameters
  
```{r}
library(datasets)
hist(airquality$Ozone)
with(airquality, plot(Wind, Ozone))
airquality <- transform(airquality, Month = factor(Month))
boxplot(Ozone ~ Month, airquality, xlab = "Month", ylab = "Ozone (ppb)", col = "snow1")
```
# Some Important Base Grphics Parameters
* pch: the plotting symbol (default is open circle): can tale numbers or characters
* lty: the line type (default is solid line), can be dashed, dotted, etc.
* lwd: the line width, specified as an integer multiple
* col: the plotting color, specified as a number, string, or hex code; The colors() function gives you a vector of colors by name
* xlab: character string for the x-axis label
* ylab: character string for the y-axis label

# par() function is used to specify global graphics parameters that affect all plots in an R session
* las: the orientation of the axis labels on the plot
* bg: the background color
* mar: the margin size
* oma: the outer margin size (default is 0 for all slides)
* mfrow: number of plots per row, column (plots are filled row-wise)
* mfcol: number of plots per row, column (plots are filled column-wise)

```{r have a look of default setting}
par("las")
par("mar")
par("mfrow")
```

# Base Plotting Functions
* plot: make a scatterplot, or other type of plot depending on the class of the object being plotted
* lines: add lines to a plot, given a vector x values and a corresponding vector of y values (or a 2 column matrix); this fnction just connects the dots;
* points: add points to a plot
* text: add text labels to a plot using specified x, y coordinates
* title: add annotations to x, y axis labels, title, subtitle, outer margin
* mtext: add arbitrary text to the margins (inner or outer) of the plot
* axis: add axis ticks/labels

```{r}
library(datasets)
with(airquality, plot(Wind, Ozone)); title(main = "Ozone and Wind in New York City") #the title function must be done with the plotting window shown, otherwise there's an error of plot.new not being called; it's only the issue in R markdown, without problem in R script

#alternative
with(airquality, plot(Wind, Ozone, main = "Ozone and Wind in New York City")); with(subset(airquality, Month == 5), points(Wind, Ozone, col = "blue")); with(subset(airquality, Month != 5), points(Wind, Ozone, col = "red")); legend("topright", pch = 1, col = c("blue", "red"), legend = c("May", "Other Months"))

#add regression model
model <- lm(Ozone ~ Wind, airquality)
with(airquality, plot(Wind, Ozone, main = "Ozone and Wind in New York City", pch = 20, abline(model, lwd = 2)))

#Multiple Base Plots
par(mfrow = c(1, 2), mar = c(4,4,2,1), oma = c(0,0,2,0)) ; with(airquality, {
  plot(Wind, Ozone, main = "Ozone and Wind")
  plot(Solar.R, Ozone, main = "Ozone and Solar Radiation")
  mtext("Ozone and Weather in New York City", outer = TRUE) ##add more space for margin and add a title to cover all plots
})


```

# Summary
* Plots in the base plotting system are created by calling successive R functions to "buil up" a plot;
* Plotting occurs in two stages:  
-- Creating of a plot  
-- Annotation of a plot (adding lines, points, text, legends)
* The base plotting system is very flexible and offers a high degree of control over plotting