---
title: "Exploratory Graphs"
author: "Yuling"
date: "4/11/2022"
output: html_document
---


* To understand data properties
* To find patterns in data
* To suggest modeling strategies
* To "debug" analyses
* To communicate results

```{r setup, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

```{r example, echo=FALSE}
#I don't have the data, only want to take a note of the reading in function parametersl
pollution <- read.csv("data/avgom25.csv", colClasses = c("numeric", "character", "factor", "numeric", "numeric"))
```

# Simple summaries of Data

## One dimension
* Five-number summary
```{r ecjo=FALSE}
#including mean, it's six number summary
summary(pollution$pm25)
```

* Boxplots
```{r}
boxplot(pollutions$pm25, col = "blue")
abline(h = 12) #in this case, it's the air quality standard
```

* Histograms
```{r}
hist(pollution$pm25, col = "green") 
rug(pollution$pm25) #detail of outliers
abline(v = 12, lwd = 2)
abline(v = median(pollution$pm25), col = "magenta", lwd = 4)
#alternative
hist(pollution$pm25, col = "green", breaks = 100) #it will change the shape of the histogram, should play around with the break number
```

* Density plot

* Barplot
+ Barplot is a good example for catagory data
```{r}
barplot(table(pollution$region), col = "wheat", main = "Number of Counties in Each Region") #main is the title of the graph
```

## Two dimensions
* Multiple/overlayed 1-D plots (Lattice/ggplot2)
```{r }
#multiple Boxplots
boxplot(pm25 ~ region, data = pollution, col = "red")

#mutiple histograms
par(mfrow = c(2,1), mar = c(4,4,2,1))
hist(subset(pollution, region == "east")$pm25, col = "green")
hist(subset(pollution, region == "west")$pm25, col = "green")
```

* Scatterplots
```{r}
with(pollution, plot(latitude, pm25))
#add color to add dimensions
with(pollution, plot(latitude, pm25, col = region))
abline(h = 12, lwd = 2, lty = 2)

#multiple scatterplots
par(mfrow = c(2,1), mar = c(5,4,2,1))
with(subset(pollution, region == "west"), plot(latitude, pm25, main = "west"))
with(subset(pollution, region == "east"), plot(latitude, pm25, main = "east"))
```

* Smooth scatterplots

## >2 dimensions
* Overlayed/multiple 2-D plots; coplots;
Use color, size, shape to add dimensions;
* Spinning plots;
* Actual 3-D plots (not that useful)

# Further resources
* [R Graph Gallery](https://r-graph-gallery.com/)
* [R Bloggers](https://www.r-bloggers.com/)

