---
title: "ggplot"
output: html_document
date: '2022-04-23'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The Basics: qplot()
* Plots are made up of aesthetics (size, shape, color) and geoms (points, lines)
* Factors are important for indicating subsets of the data (if they are to have different properties); they should be labeled
* The qplot() hides detials underneath, which is okay for most operations
* ggplot() is the core function and very flexible for doing things qplot() cannot do

```{r Example dataset}
library(ggplot2)
str(mpg)
qplot(x = displ, y = hwy, data = mpg, color = drv)

## adding a geom (a summary of the data)
qplot(x = displ, y = hwy, data = mpg, geom = c("point", "smooth"))

##histograms
qplot(hwy, data = mpg, fill = drv)

## Facets (like panels)
qplot(displ, hwy, data = mpg, facets = .~drv) #notice the position of "~." or ".~", the former of which splits coloumns and the latter one splits rows
qplot(hwy, data = mpg, facets = drv~., binwidth = 2)
```

# Basic Components of a ggplot2 Plot
* A data frame
* aesthetic mappings: how data are mapped to color, size
* geoms: geometric objects like points, lines, shapes
* facets: for conditional plots
* stats: statistical transformations like binning, quantiles, smoothing.
* scales: what scale an aesthetic map uses (example: male = red, female = blue)
* coordinate system

# Building Plots with ggplot2
* When building plots in ggplot2 (rather than using qplot) the "artist's palette" model may be the closest analogy
* Plots are built up in layers
-- plot the data
-- overlay a summary
-- metadata and annotation

```{r eval=FALSE}
# There's no data frame for the codes, don't run it
g <- ggplot(maacs, aes(logpm25, NocturnalSympt))
summary(g)
p <- g + geom_point()
print(p) # can also do without assign the plotting to "p"
# Add more layers
g + geom_point() + geom_smooth(method = "lm") #add Smooth via linear regression line
# Add more layers: Facets
g + geom_point() + facet_grid(.~bmicat) + geom_smooth(method = "lm") #labels are automatically set on plotting, therefore, it's better practice to always give informative labels in factors of the data frame being plotted

```

# Annotation
* Labels: xlab(), ylab(), labs(), ggtile()
* Each of the "geom" functions has options to modify
* For things that only make sense globally, use theme()  
  -- Example: theme(legend.position = "none")
* Two standard appearance themes are included   
  -- theme_grey(): The default theme (grey background)   
  -- theme_bw(): More stark-plain  
  
```{r eval=FALSE}
# Modifying Aesthetics
g + geom_point(color = "steelblue", size = 4, alpha = 1/2) # add constant color values for whole plotting
g + geom_point(aes(color = bmicat), size = 4, alpha = 1/2) # add colors on data variable
g + geom_point(aes(color = bmicat)) + labs(title = "MAACS Cohort") + labs(x = expression("log " * PM[2.5]), y = "Nocturnal Symptoms")

# change the theme
g + geom_point(aes(color = bmicat)) + theme_bw(base_family = "Times")
```

# A note about axis limits
* it depends on whether the outlier is interesting for the research
```{r}
testdat <- data.frame(x = 1:100, y = rnorm(100))
testdat[50,2] <- 100 ##outlier
plot(testdat$x, testdat$y, type = "l", ylim = c(-3, 3)) #set the y-limits to conceal outliers
g <- ggplot(testdat, aes(x = x, y = y))
g + geom_line() #by default, the outlier will be included
g + geom_line() + ylim(-3, 3) #by setting axis limits, ggplot will subset data within the region
g + geom_line() + coord_cartesian(ylim = c(-3, 3)) # outlier included
```


# cut continious variable into tertiles to work as conditional factor

```{r eval=FALSE}
## caculate the deciles of the data
cutpoints <- quantile(maacs$logno2_new, seq(0,1,length = 4), na.rm = T)
## cut the data at the deciles and create a new factor variable
maacs$no2dec <- cut(maacs$logno2_new, cutpoints)
## see the levels of the newly created factor variable
levels(maacs$no2dec)

# example codes without data frame
g <- ggplot(maacs, aes(logpm25, NocturnalSymt))

## add layers, it's good practice to assign a new object for it
g + geom_point(alpha = 1/3)# add points
  + facet_wrap(bimcat ~ no2dec, nrow = 2, ncol = 4) # make panels
  + geom_smooth(method = "lm", se = FALSE, col = "steelblue") # add smoother
  + theme_bw(base_family = "Avenir", base_size = 10) # change theme
  + labs(x = expression("log " * PM[2.5]))
  + labs(y = "Nocturnal Symptoms")
  + labs(title = "MAACS Cohort")

```
