---
title: "Hierarchical clustering"
author: "Yuling"
date: "4/26/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Hierarchical clustering
* An agglomerative approach  
  -- Find closest two things
  -- Put them together
  -- Find next closest
  
* Requires  
  -- A defined distance
  -- A merging approach
  
* Produces  
  -- A tree showing how close things are to each other
  
  
# Define "close"
* Distance or similarity  
  -- Continuous - euclidean distance
  -- Continuous - correlation similarity
  -- Binary - manhattan distance
  
* Pick a distance/similarity that makes sense for your problem

# Hierarchical clustering - example
```{r}
set.seed(1234) # make the random data reproducible
par(mar = c(0,0,0,0))
x <- rnorm(12, mean = rep(1:3, each = 4), sd = 0.2)
y <- rnorm(12, mean = rep(c(1,2,1), each = 4), sd = 0.2)
plot(x, y, col = "blue", pch = 19, cex = 2); text(x + 0.05, y + 0.05, labels = as.character(1:12))

dataFrame <- data.frame(x = x, y = y)
distxy <- dist(dataFrame) #serves as default for euclidean distance
hClustering <- hclust(distxy)
plot(hClustering)
```
[RGraphGallery](https://r-graph-gallery.com/)

# heatmap()
```{r}
dataFrame <- data.frame(x = x, y = y)
set.seed(143)
dataMatrix <- as.matrix(dataFrame)[sample(1:12), ]
heatmap(dataMatrix)
```

