---
title: "Lattice Plotting System"
author: "Yuling"
date: "4/22/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The Lattice Plotting System
## Two packages:
* Lattice: contains code for producing Trellis graphics, which are independent of the "base" graphics system; includes functionss like xyplot, bwplot, levelplot;
* grid: implements a different graphing system independent of the "base" system; the lattice package builds on top of grid;

-- We seldom call functions from the grid package directly;

* The lattice plotting system does not have a "two-phase" aspect with separate plotting and annotation like in base plotting;
* All plotting/annotation is done at once with a single function call

# Lattice Functions
* xyplot: create scatterplots
```{r } 
# xyplot(y ~ x | f * g, data)
# # f and g are conditioning variables - they are optional
# -- the * indicates an interaction between two variables

library(lattice)
library(datasets)
xyplot(Ozone ~ Wind, data = airquality)

# Convert "Month" to a factor variable
airquality <- transform(airquality, Month = factor(Month))
xyplot(Ozone ~ Wind | Month, data = airquality, layout = c(5,1))
```

* bwplot: box-and-whiskers plots ("boxplots")
* histogram: histograms
* stripplot: like a boxplot but with actual points
* splom: scatter matrix; like pairs in base plotting system
* levelplot, contourplot: for plotting "image" data

# Lattice Behaviour
* Base graphics functions plot data directly to the graphic device (screen, PDF file, etc);
* Lattice graphics functions return an object of class trellis
* The print methods for lattice functions actually do the work of plotting the data on the graphics device
* Lattice functions return "plot objects" that can, in principle, be stored (but it's usually better to just save the code + data).
* On the command line, trellis objects are auto-printed so that it appeats the function is plotting the data

```{r}
# compared with auto-printing
p <- xyplot(Ozone ~ Wind, data = airquality)
print(p)
```

# Lattice Panel Functions
* Lattice functions have a __panel function__ which controls what happens inside each panel of the plot
* Panel function receive the x/y coordinates of the data points in their panel (along with any optional arguments)

```{r}
set.seed(10)
x <- rnorm(100)
f <- rep(0:1, each = 50)
y <- x + f - f * x + rnorm(100, sd = 0.5)
f <- factor(f, labels = c("Group 1", "Group 2"))
xyplot(y ~ x | f, layout = c(2,1))

# Custom panel function
xyplot(y ~ x | f, panel = function(x, y, ...) {
  panel.xyplot(x, y, ...) # first call the default panel function for "xyplot"
  panel.abline(h = median(y), lty = 2) ## add a horizontal line at the median
})

# add a simple linear regression line
xyplot(y ~ x | f, panel = function(x, y, ...) {
  panel.xyplot(x, y, ...) # first call the default panel function for "xyplot"
  panel.lmline(x, y, col = 2) ## add a horizontal line at the median
})
```

# Summay
* Lattice plots are constructed with a single function call to a core lattice function (e.g. xyplot)
* Aspects like margins and spacing are automatically handled and defaults are usually sufficient
* The lattice system is ideal for creating conditioning plots where you examine the same kind of plot under many different conditions
* Panel functions can be specific/ customized to modify what is plotted in each of the plot panels

