---
title: "Plotting Systems in R"
author: "Yuling"
date: "4/12/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# The Base Plotting System
* "Artist's palette" model
* Start with blank canvas and build up from there
* Start with **plot function** (or similar)
* Use **annotation functions** to add/modify (text, lines, points, axis)

## Drawbacks
* Can't go back once plot has started; need to plan in advance
* no conceptual language

```{r Base plot}
library(datasets)
data(cars)
with(cars, plot(speed, dist))
```

# The Lattice system
* Plots are created with a single functional call (xyplot, bwplot, etc.)
* Most useful for conditionaing types of plots: Looking at how y changes with x across levels of z
* Things like margins/spacing set automatically because entire plot is specified at once
* Good for putting many plots on a screen

## Drawbacks
* Difficult to specify an entire plot in a single function call
* Annotation in plot is not especially intuitive
* Use of panel functions and subscripts difficult to wield and requires intense preparation
* Canoot "add" to the plot once it is created

```{r}
library(lattice)
state <- data.frame(state.x77, region = state.region)
xyplot(Life.Exp ~ Income | region, data =state, layout = c(4,1))
```

# The ggplot2 System
* Splits the difference between base and lattice in a number of ways
* Automatically deals with spacing, text, titles but also allows you to annotate by "adding" to a plot
* Superficial similarity to lattice but gereally easier/more intuitive to usel
* Default mode makes many choices for you (but you can still customize to your heart's desire)

```{r}
library(ggplot2)
data(mpg)
qplot(displ, hwy, data = mpg)
```

